%starting file for literature review
%starting file for Chapter 2 CD105 cruise remove later for insertion
\documentclass[12pt,twoside]{mythesis}
%,draft
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{subfigure}
\usepackage{float}
\usepackage[section]{placeins} % for preventing floats drifting
\usepackage{sidecap}
\usepackage[font=small]{caption} % for smaller captions
\usepackage{amssymb} % for maths
% for fancy chapter headings
\usepackage{psboxit,pstcol}
\input{mydefs}

\title{The Coastal Transition Zone of the Galician Region: Upper Layer Response to Wind Forcing and Circulation}
\author{Ricardo Torres}
%
\month{\textsc{May}} \year{2003}
\previousdegrees{B.Sc., Universidad de Vigo, Spain, 1996\\%
  M.Sc. (Hons), School of Ocean Sciences, Bangor University, 1998}
\degreetitle{Doctor of Philosophy}
\department{Physical Oceanography}
% \author{}{Torres, Ricardo}
% \degree{Doctor of Philosophy}{Ph.D.}{May}{2003}

%
\includeonly{/home/rito/Research/PhD/final_2023/literature/literature,
/home/rito/Research/PhD/final_2023/chapter2/cd105,/home/rito/Research/PhD/final_2023/chapter1/winds,
/home/rito/Research/PhD/final_2023/chapter3/cd114,/home/rito/Research/PhD/final_2023/chapter4/winter,%
/home/rito/Research/PhD/final_2023/chapter5/drifters,/home/rito/Research/PhD/final_2023/appendix/fly, %
/home/rito/Research/PhD/final_2023/chapter6/conclusions,introduction}

\begin{document}

\maketitle

\chapter*{Abstract}

\chaptermark{abstract}
\addcontentsline{toc}{chapter}{Abstract}

Through a set of observations including satellite, cruise and
mooring data encompassing the period 1997-2001 the Galician shelf
response to the varying seasonal forcing  has been  characterised.
The typical spatial variability of the wind and its effect on the
shelf circulation has been investigated. The seasonality of the
large scale wind during July 1999- May 2001 was masked by
upwelling/downwelling patterns throughout the year. However,
upwelling winds were found to be more consistent during the summer
and the system showed a clear seasonality with upwelling during
the summer and downwelling during winter mediated by a meridional
density gradient. Despite its spatial complexity, the mesoscale
wind variability was reduced to a finite number of spatial
patterns. These favoured upwelling North or South of Cape
Finisterre or downwelling along the west coast of Galicia. Their
relative abundance determined the interannual variability during
both upwelling and downwelling regimes.

Slope poleward flow characteristic of the downwelling regime was
measured in all regimes, suggesting it is a permanent feature in
the Galician CTZ system. Its winter variability was related to
unusual periods of upwelling winds. It had a surface signature in
winter, autumn and during the spring. In the summer it forms a
deep poleward flow. The shelf is effectively isolated from the
ocean due to the shelf-edge front as indicated by the lagrangian
horizontal diffusivity although some exchange still takes place in
the form of a bottom Ekman layer.

The weakening of the meridional density gradients sparks the decay
phase of the poleward slope current. The warm tongue broadens, and
is distorted by interactions with the mesoscale eddy field and
freshwater inputs. Upwelling favourable winds push the flow
offshore establishing a second northward flow, and short lived
eddies and upwelling dynamics break the slope flow structure.

During the upwelling regime, the upper column structure becomes
dominated by coastal upwelling and filaments and the poleward flow
is restricted to deeper levels. A detailed survey of a filament
showed limited export capabilities, although the lagrangian
diffusivity was enhanced during the summer upwelling regime.
Interannual variability during the upwelling regime is
characterised by either presence/absence of filaments or whether
the main filament is at Cape Finisterre or at 42\deg latitude.

Towards the end of the upwelling regime, when the upwelling winds
start weakening, the poleward flow climbs back onto the upper
slope. However, it is not until the meridional density strengthens
that the poleward flow surfaces again. If filaments are still
present, the slope poleward flow cuts their water supply and they
slowly mix away.
% \end{abstract}
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to express my gratitude to my supervisor Dr. Des
Barton for the invaluable guidance and support during the entire
gestation of this thesis.

Part of this work was written during a stay in the CICESE station
of La Paz, Mexico, and useful discussions with Dr. Trasvi\~na and
Dr. Gutierrez helped in the development of the work.

Many of the data presented here were collected during four cruises
carried out under the OMEX II-II EU project and the crew, officers
and scientific personnel made it both possible and highly
enjoyable. My gratitude to BODC for their fantastic work in the
quality control of the cruise data.

Thanks to the RSDAS team, Plymouth Marine Laboratory for supplying
the remote sensing data.

These years in Menai Bridge have been shared and enjoyed with a
great bunch of friends who supported, helped and made life so much
better; my brother Nacho, Teri, Maite, Nuno, Fernando, Jose, Mark
and many more. Thanks for the constant source of support and
distraction!

% My deepest and warmest thanks go to Dr. Loopy, my one and only
% source of inspiration.

\allcontents
% \singlespacing
% \hyphenpenalty=10000
% \tableofcontents
% \listoftables
% \listoffigures
% \hyphenpenalty=3000
% \textspacing



\cleardoublepage
\setcounter{chapter}{0}
\mainbody

%\begin{dedication}
%A mis padre que estan deseando como locos que me vuelva y no se
%cuando podre darles esa alegria.
%\end{dedication}

%
% \setlength{\parindent}{0pt}                     % after table of contents
% \setlength{\parskip}{0.5cm}% so it doesn't fuck it up
% %1ex plus 0.5ex minus 0.7ex
% \setlength{\textheight}{9.25truein}             % doesn't indent subsections

\input{/home/rito/Research/PhD/final_2023/introduction}

% For enabling pdflatex
\DeclareGraphicsExtensions{.png}

\graphicspath{{/home/rito/Research/PhD/final_2023/Literature/graphics/}}
\input{/home/rito/Research/PhD/final_2023/Literature/literature}
%
\graphicspath{{/home/rito/Research/PhD/final_2023/chapter1/graphics/}}
\input{/home/rito/Research/PhD/final_2023/chapter1/winds}

\graphicspath{{/home/rito/Research/PhD/final_2023/chapter2/graphics/}}
\input{/home/rito/Research/PhD/final_2023/chapter2/cd105}

\graphicspath{{/home/rito/Research/PhD/final_2023/chapter3/graphics/}}
\input{/home/rito/Research/PhD/final_2023/chapter3/cd114}
%
\graphicspath{{/home/rito/Research/PhD/final_2023/chapter4/graphics/}}
\input{/home/rito/Research/PhD/final_2023/chapter4/winter}
%
\graphicspath{{/home/rito/Research/PhD/final_2023/chapter5/graphics/}}
\input{/home/rito/Research/PhD/final_2023/chapter5/drifters}

\graphicspath{{/home/rito/Research/PhD/final_2023/chapter6/graphics/}}
\input{/home/rito/Research/PhD/final_2023/chapter6/conclusions}

\formatbibliography
\bibliographystyle{agufull}
\bibliography{/home/rito/Research/PhD/final_2023/references/literature,/home/rito/Research/PhD/final_2023/references/poleward,/home/rito/Research/PhD/final_2023/references/drifters_refs}
% \setlength{\parindent}{0pt}                     % after table of contents
% \setlength{\parskip}{1ex plus 0.5ex minus 0.7ex}% so it doesn't fuck it up
% \setlength{\textheight}{9.25truein}             % doesn't indent subsections


\formatappendices
\input{/home/rito/Research/PhD/final_2023/appendix/fly}

%
\end{document}
